---
output: pdf_document
title: CON Data Analysis
author: Will St. John
date: April 30, 2024
---

# Libraries
This code loads in the required r libraries to streamline the data reduction process.
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(googlesheets4)
library(ggplot2)
```

# Data
In this section, I use the `googlesheets4` library to read in the tidy data from the Google Sheet.
```{r, warning=FALSE, message=FALSE}
# Read the data from the Google Sheet
df <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1SOdDMR_2ekxp_qxyp4vmCBdVigUE3S7GMWIfGzz4r2E/edit#gid=356557056", sheet = "TidyData")
```

# Data Cleaning
A new column of average masses from the three mass columns (`LogM_U`, `LogM_Leroy`, and `LogM_Other`) is made. A summary of the data is then printed to the console.
```{r, warning=FALSE, message=FALSE, results = 'hide' }
# Data Cleaning
df$LogM <- rowMeans(df[, c("LogM_U", "LogM_Leroy", "LogM_Other")], na.rm = TRUE)
summary(df)

df$L_Akylas <- df$`L_14-195keV_Akylas` * 1e42 / 3.826e33 / 1e11
df$type <- ifelse(df$CONS == TRUE, "CON", "LIRG")
df$type <- ifelse(df$CT_AGN_Akylas == TRUE, "CT_AGN", df$type)

df$L <- ifelse(!is.na(df$L_IR_Falstad), df$L_IR_Falstad, NA)
df$L <- ifelse(!is.na(df$L_Akylas), df$L_Akylas, df$L)


colnames(df)
df$BeamSize_Falstad
df$BeamSize_Falstad <- strsplit(df$BeamSize_Falstad, "x")
df$BeamSize_Falstad <- sapply(
        df$BeamSize_Falstad,
        function(x) as.numeric(x[1]) * as.numeric(x[2]))
df$BeamSize_Falstad <- sqrt(df$BeamSize_Falstad) * df$Hubble_Distance_Mpc / 206265
df$BeamSize_Falstad
```

# Data Analysis
## Bias Comparison
```{r, warning=FALSE, message=FALSE}
df %>% filter(!is.na(BeamSize_Falstad)) %>%
        ggplot(aes(x = Hubble_Distance_Mpc / 1e6, y = BeamSize_Falstad / 1e6, color = CONS)) +
        geom_point(size = 3) +
        ggtitle("Beam Size [pc] vs Distance [pc]") +
        xlab("Hubble Distance [pc]") +
        ylab("Linear Beam Size Proxy [pc]")

df %>% filter(!is.na(BeamSize_Falstad)) %>%
        ggplot(aes(x = L_IR_Falstad, y = BeamSize_Falstad / 1e6, color = CONS)) +
        geom_point(size = 3) +
        ggtitle("Beam Size [pc] vs Luminosity [10^11 L_sun]") + 
        xlab("Luminosity [10^11 L_sun]") +
        ylab("Linear Beam Size Proxy [pc]")
```


```{r, warning=FALSE, message=FALSE}
df %>% ggplot(aes(x = Hubble_Distance_Mpc, y = L, color = type)) +
        geom_point(size = 3) +
        ggtitle("Luminosity (L) vs. Hubble Distance (Hubble_Distance_Mpc)")
```

```{r, warning=FALSE, message=FALSE}
df %>% ggplot(aes(x = Hubble_Distance_Mpc, color = type, fill = type)) +
        # geom_histogram(alpha = 0.5) +
        geom_density(alpha = 0.5) +
        ggtitle("Hubble Distance Density Plot")
```

```{r, warning=FALSE, message=FALSE}
df %>% ggplot(aes(x = L, color = type, fill = type)) +
        # geom_histogram(alpha = 0.5) +
        geom_density(alpha = 0.5) +
        ggtitle("Luminosity Density Plot") +
        coord_trans(x = "log10")

```

<!-- 
## Scatter Plot
In this section, I create bivariate data visualizations for different variables in the data set. The first is a scatter plot of the log of the average mass (`LogM`) and the Hubble distance in Mpc (`Hubble_Distance_Mpc`). 

```{r, warning=FALSE, message=FALSE, include=FALSE}
# Scatter Plot
df %>% ggplot(
        aes(x = LogM,
        y = Hubble_Distance_Mpc,
        color = CONS)) +
        geom_point() +
        ggtitle("Log of Average Mass (LogM) vs. Hubble Distance (Hubble_Distance_Mpc)")
```

The second is a scatter plot of the log of the average mass (`LogM`) and the Falstad et al. distance in Mpc (`D_L_Falstad`) with the size of the points representing the infrared luminosity in solar luminosities (`L_IR_Falstad`). 
```{r, warning=FALSE, message=FALSE, include=FALSE}
df2 <- df %>%
        filter(!is.na(LogM) & !is.na(D_L_Falstad)) %>%
        select(LogM, D_L_Falstad, L_IR_Falstad, CONS)

df2 %>% ggplot(
        aes(x = LogM,
        y = D_L_Falstad,
        color = CONS, size = L_IR_Falstad)) +
        geom_point() +
        ggtitle("Log of Average Mass (LogM) vs. Falstad Distance (D_L_Falstad)")
```

The third is a scatter plot of the log of the average mass (`LogM`) and the infrared luminosity in solar luminosities (`L_IR_Falstad`) with the size of the points representing the Falstad distance in Mpc (`D_L_Falstad`).
```{r, warning=FALSE, message=FALSE, include=FALSE}
df2 %>% ggplot(
        aes(x = LogM,
        y = L_IR_Falstad,
        color = CONS, size = D_L_Falstad)) +
        geom_point() +
        ggtitle("Log of Average Mass (LogM) vs. Infrared Luminosity (L_IR_Falstad)")
```

## Density Plot and Histogram
In this section, I create univariate data visualizations for different variables in the data. Specifically, I show density plots and histograms for the log of the average mass (`LogM`) for the entire data ...

```{r, warning=FALSE, message=FALSE, include=FALSE}
# Density Plot and Histogram
df %>% ggplot(
        aes(x = LogM, color = CONS, fill = CONS)) +
        geom_density(alpha = 0.5) +
        ggtitle("Log of Average Mass (LogM)") +
        geom_histogram(alpha = 0.2) # optional
```

... the log of the average mass (`LogM`) for the data from the Falstad et al paper ...

```{r, warning=FALSE, message=FALSE, include=FALSE}
df2 %>% ggplot(
        aes(x = LogM, color = CONS, fill = CONS)) +
        geom_density(alpha = 0.5) +
        ggtitle("Log of Average Mass (LogM) for Falstad et al. paper") +
        geom_histogram(alpha = 0.2) # optional
```
... the Falstad et al. distance in Mpc (`D_L_Falstad`) ...
```{r, warning=FALSE, message=FALSE, include=FALSE}
df2 %>% ggplot(
        aes(x = D_L_Falstad, color = CONS, fill = CONS, title = "test")) +
        geom_density(alpha = 0.5) +
        ggtitle("Falstad Distance (D_L_Falstad)")
        # geom_histogram(alpha = 0.2) # optional
```
... and the infrared luminosity in solar luminosities (`L_IR_Falstad`) from the Falstad paper.

```{r, warning=FALSE, message=FALSE, include=FALSE}
df2 %>% ggplot(
        aes(x = L_IR_Falstad, color = CONS, fill = CONS)) +
        geom_density(alpha = 0.5) +
        ggtitle("Infrared Luminosity (L_IR_Falstad)") +
        geom_histogram(alpha = 0.2) # optional
```


## Hierarchical Clustering
```{r, warning=FALSE, message=FALSE, include=FALSE}
# Hierarchical Clustering
df_good <- df %>%
        filter(!is.na(LogM) & !is.na(Hubble_Distance_Mpc)) %>%
        select(Galaxy, LogM, Hubble_Distance_Mpc, CONS)

df_dist_mass <- df_good %>%
        select(LogM, Hubble_Distance_Mpc)


cluster_complete <- hclust(dist(df_dist_mass), method = "complete")
cluster_single <- hclust(dist(df_dist_mass), method = "single")
cluster_average <- hclust(dist(df_dist_mass), method = "average")
cluster_centroid <- hclust(dist(df_dist_mass), method = "centroid")

plot(cluster_complete, labels = as.character(df_good$CONS))
plot(cluster_single, labels = as.character(df_good$CONS))
plot(cluster_average, labels = as.character(df_good$CONS))
plot(cluster_centroid, labels = as.character(df_good$CONS))
```

## Partitioning Around Medoids (PAM)
```{r, warning=FALSE, message=FALSE, include=FALSE}
## Partitioning Around Medoids (PAM)
pam_gal <- cluster::pam(df_dist_mass, k = 6)
summary(pam_gal)
plot(pam_gal)
pam_gal$medoids

df %>% ggplot(
        aes(x = LogM,
        y = Hubble_Distance_Mpc,
        color = CONS)) +
        geom_point() +
        geom_point(
                data = as.data.frame(pam_gal$medoids),
                aes(x = LogM, y = Hubble_Distance_Mpc),
                color = "black",
                size = 3)
```

 -->


